---
title: "Take-home Exercise 2"
author: "Jeffery Lau"
date: 09-23-2024
date-modified: "last-modified"
description: |
  In this exercise, we will be analyzing drug use offence in Thailand from 2017 - 2022, using spatial autocorrelation techniques to identify clusters, outliers and hotspots of such offence. We'll also create maps to visualize the results and interpret the spatial patterns.
categories:
  - Take-home
format:
  html:
    toc: true
execute: 
  eval: true
  echo: true
  warning: false
  freeze: false
---

# 1. Introduction

## 1.1 Context

Drug abuse is a significant global health issue with far-reaching consequences, including negative health outcomes, financial burdens, and social problems. Despite efforts to combat it, illicit drug consumption remains prevalent and continues to grow. In 2021, an estimated 296 million people aged 15–64 worldwide had used a drug in the past 12 months.

Thailand, situated near the Golden Triangle - the largest drug production site in Asia - has become a market and transit route for drug trafficking due to its geographic location and infrastructure development. This has made drug abuse a major social issue within the country.

Youth are particularly vulnerable to drug abuse, with approximately 2.7 million Thai youths using drugs. Among 15-19-year-olds, around 300,000 require drug treatment. Vocational-school students are disproportionately affected, with nearly double the number of drug users compared to secondary-school students.

## 1.2 Objectives

1.  Determine if drug abuse indicator are spatially dependent
2.  Identify clusters, outliers and hotspots of drug abuse 
3.  Analyse how the spatial patterns

# 2. Setup

## 2.1 Loading Packages

In this project we will be using the following packages:

-   **`tmap`:** A package for creating thematic maps in R. It provides a simple and flexible interface for customizing maps and exporting them in various formats.

-   **`sf`:** A package for simple feature data, which is a modern standard for representing geographic features. It provides efficient data structures and functions for spatial operations.

-   **`sfdep`:** A collection of R packages designed for data manipulation, visualization, and analysis. It includes packages like dplyr, ggplot2, and tidyr, which are commonly used in data science workflows.

-   **`raster`:** A package for working with raster data, such as images and digital elevation models. It offers tools for reading, writing, manipulating, and analyzing raster data.

-   **`spatstat`:** A package for spatial point pattern analysis. It provides functions for analyzing the distribution and relationships between points in a study area.

-   **`maptools`:** A package for reading and writing various spatial data formats, including shapefiles and KML. It also provides tools for converting between different spatial data structures.

-   **`sp`:** A package for spatial data in R. It provides classes and functions for representing and manipulating spatial data, such as points, lines, and polygons.

-   **`tidyverse`:** A collection of R packages designed for data manipulation, visualization, and analysis. It includes packages like dplyr, ggplot2, and tidyr, which are commonly used in data science workflows.

```{r}
pacman::p_load(tmap, sf, sfdep, raster, spatstat, maptools, tidyverse, DT)
```

## 2.2 Loading Data

The dataset used in this analysis is sourced from the Armed Conflict Location & Event Data (ACLED) database. It contains information on various types of armed conflict events, including Battles, Explosion/Remote violence, Strategic developments, Protest, Riots, and Violence against civilians. The data covers the period from January 2021 to June 2024 and provides specific details such as event date, location coordinates, actor information, and more.

```{r}
off_sf <- read_csv("data/aspatial/thai_drug_offenses_2017_2022.csv")
```

Here we will be using the read_sf as the st_read encountered 

```{r}
prov_sf <- read_sf('data/geospatial/tha_adm_rtsd_itos_20210121_shp/tha_admbnda_adm1_rtsd_20220121.shp')
```


# 3. Data Wrangling

## 3.1 Examining Data 

Before performing the join we need to identify if the column we are joining share the same province spelling.

```{r}
prov_diff <- function(off_sf_provinces, prov_sf_provinces) {
  # Check if all provinces from off_sf are in prov_sf
  all_off_sf_in_prov_sf <- all(off_sf_provinces %in% prov_sf_provinces)
  
  # Check if all provinces from prov_sf are in off_sf
  all_prov_sf_in_off_sf <- all(prov_sf_provinces %in% off_sf_provinces)
  
  # Check if the sets are equal
  sets_are_equal <- all_off_sf_in_prov_sf && all_prov_sf_in_off_sf
  
  if (sets_are_equal) {
    cat("All province names match.\n")
  } else {
    cat("There are mismatches in province names.\n")
  
    # Find the differences
    cat("Missing in prov_sf:", paste(setdiff(off_sf_provinces, prov_sf_provinces), collapse = ", "), "\n")
    cat("Missing in off_sf:", paste(setdiff(prov_sf_provinces, off_sf_provinces), collapse = ", "), "\n")
  }
}

prov_diff(unique(off_sf$province_en), unique(prov_sf$ADM1_EN))
```

Notice in the above when we compare the unique set from both SF there is 2 misspellings, where we will be renaming the province in off_sf:
1.    Loburi -> Lop Buri
2.    Buogkan -> Bueng Kan

```{r}
off_sf <- off_sf %>%
  mutate(province_en = case_when(
    province_en == "Loburi" ~ "Lop Buri",
    TRUE ~ province_en
  ))
off_sf <- off_sf %>%
  mutate(province_en = case_when(
    province_en == "buogkan" ~ "Bueng Kan",
    TRUE ~ province_en
  ))

prov_diff(unique(off_sf$province_en), unique(prov_sf$ADM1_EN))
```

## 3.2 Joining `sf` on Province

```{r}
du_sf <- left_join(off_sf, prov_sf, by = c("province_en" = "ADM1_EN"))
head(du_sf)
```

## 3.3 Data Cleaning

### 3.3.1 Detecting NA

```{r}
# Check for null values in key columns
na_count <- du_sf %>%
  summarise(na_geometry = sum(is.na(geometry)),
            na_province = sum(is.na(province_en)),
            na_drug_offense = sum(is.na(types_of_drug_offenses)),
            na_cases = sum(is.na(no_cases)))

print(na_count)
```

### 3.3.2 Ensuring Consistency

The data should have the years 2017 - 2022 for 16 types of drug offenses:

```{r}
# Check for null values in key columns
print(unique(du_sf$fiscal_year))
print(unique(du_sf$types_of_drug_offenses))
```

Next we need to ensure all province should have 96 records each. We can also verify that since there 77 unique province which tallies to 7392 obs. in our `sf` given each province have 96 records

```{r}
# Assuming du_sf is a data frame
province_counts <- du_sf %>%
  group_by(province_en) %>%
  summarize(count = n())

# Check if all counts are 96
all_96 <- all(province_counts$count == 96)

if (all_96) {
  cat("All provinces have 96 records.\n")
} else {
  cat("Provinces with counts other than 96:\n")
  print(province_counts %>% filter(count != 96))
}
```

### 3.3.3 Detecting Duplicates

Here we try to detect if there happen to be any duplicate given a state, fiscal year and typeo of offence.

```{r}
duplicates <- du_sf %>%
  group_by(province_en, fiscal_year, types_of_drug_offenses) %>%
  filter(n() > 1)

if (nrow(duplicates) > 0) {
  cat("Duplicates found:\n")
  print(duplicates)
} else {
  cat("No duplicates found.\n")
}
```

# 4. Exploratory Data Analysis

```{r}

```

