---
title: "Take-home Exercise 3"
author: "Jeffery Lau"
date: 10-13-2024
date-modified: "last-modified"
description: |
  In this exercise, we will be looking prototyping for my team's Geospatial Analytics Shiny Application. I am responsible for 
categories:
  - Take-home
format:
  html:
    toc: true
execute: 
  eval: true
  echo: true
  warning: false
  freeze: false
---

# 1. Introduction

## 1.1 Responsibility

## 1.2 Motivation

## 1.3 Objectives


# 2. Packages and Data

## 2.1 Packages Required

```{r}
pacman::p_load(sf, st, tidyverse, raster, tmap, tmaptools, ggplot2, spatstat, sfdep, ClustGeo, 
               ggpubr, cluster, factoextra, NbClust,
               heatmaply, corrplot, psych, tidyverse, GGally)
```

## 2.2 Data Required

## 2.2.1 Datasets

There are two datasets being used in this exercise.

-   [Malaysia – Crime by District and Crime Type](https://data.gov.my/data-catalogue/crime_district) from data.gov.my in csv format.

-   [Malaysia - Subnational Administrative Boundaries](https://data.humdata.org/dataset/cod-ab-mys) with included administrative regions in shapefile format.

We first import the crime rate csv file into R.

```{r}
crime_df <- read_csv("data/aspatial/crime_district.csv")
```

Next, we import the administrative regions of Malaysia.

```{r}
mys_sf <- read_sf(dsn = "data/geospatial/mys_adm_unhcr_20210211_shp", 
                 layer = "mys_admbnda_adm2_unhcr_20210211") %>%
  st_as_sf(coords =c(
    "longitude", "latitude"),
           crs = 4326) %>%
  st_transform(crs = 3167)
```

## 2.2.2 Filtering

However in this case study, for ease of analysis, we choose to focus on West Malaysia, and thus will be filtering out Sarawak, Sabah and Labuan.

```{r}
mys_sf <- mys_sf %>%
          filter(ADM1_EN != 'W.P. Labuan' & ADM1_EN != 'Sabah' & ADM1_EN != 'Sarawak') %>%
          mutate(ADM1_EN = replace(ADM1_EN, ADM1_EN == 'W.P. Kuala Lumpur', 'Kuala Lumpur'),
                 ADM1_EN = replace(ADM1_EN, ADM1_EN == 'W.P. Putrajaya', 'Selangor'), 
                 ADM1_EN = toupper(ADM1_EN),
                 ADM2_EN = toupper(ADM2_EN))

mys_sf
```

```{r}
crime_df <- crime_df %>%
              filter(state != 'Malaysia' & state != 'Sabah' & state != 'Sarawak' & district != 'All') %>%
              mutate(year = year(date),
                     state = replace(state, state == 'W.P. Kuala Lumpur', 'Kuala Lumpur'),
                     state = toupper(state),
                     district = toupper(district))
crime_df
```

```{r}
# Assuming you have two character vectors:
state_crime <- unique(crime_df$state)
state_sf <- unique(mys_sf$ADM1_EN)

# Find states in crime_df that are not in mys_sf
missing_in_sf <- setdiff(state_crime, state_sf)

# Find states in mys_sf that are not in crime_df
missing_in_crime <- setdiff(state_sf, state_crime)

# Print the mismatches
print("States in crime_df not found in mys_sf:")
print(missing_in_sf)

print("States in mys_sf not found in crime_df:")
print(missing_in_crime)
```

```{r}
# Assuming you have two character vectors:

district_crime <- unique(crime_df$district)

district_sf <- unique(mys_sf$ADM2_EN)



# Find districts in crime_df that are not in mys_sf

missing_in_sf <- setdiff(district_crime, district_sf)



# Find districts in mys_sf that are not in crime_df

missing_in_crime <- setdiff(district_sf, district_crime)



# Print the mismatches

print("Districts in crime_df not found in mys_sf:")

print(missing_in_sf)



print("Districts in mys_sf not found in crime_df:")

print(missing_in_crime)
```



```{r}
# Create a new column in crime_df
crime_df <- crime_df %>%
  mutate(state_district = paste(state, district, sep = "_"))

# Create a new column in mys_sf
mys_sf <- mys_sf %>%
  mutate(state_district = paste(ADM1_EN, ADM2_EN, sep = "_"))

# Now, you can compare the unique values in the `state_district` column of both data frames:

# Assuming you have two character vectors:
state_district_crime <- unique(crime_df$state_district)
state_district_sf <- unique(mys_sf$state_district)

# Find mismatches
missing_in_sf <- setdiff(state_district_crime, state_district_sf)
missing_in_crime <- setdiff(state_district_sf, state_district_crime)

# Print the mismatches
print("State-District combinations in crime_df not found in mys_sf:")
print(missing_in_sf)

print("State-District combinations in mys_sf not found in crime_df:")
print(missing_in_crime)
```


## 2.3 Data Wrangling

### 2.3.1 Resolving data duplication

There is also an issue of double counting data, where state 'Malaysia' comprises of all the crimes, and type of crime 'all' also comprises of its type of crime.

As we are doing by state and not by district, we remove the double counting caused by district. At the same time, we will also remove Sabah and Sarawak, which are not the focus of our current analysis.

```{r}
crime_df <- crime_df %>%
                  filter(state !='Malaysia') %>% 
                  filter(type !='all') %>% 
                  filter(district =='All') %>%
                  filter(state != 'Sabah') %>%
                  filter(state != 'Sarawak') %>%
                  filter(state != 'Labuan')
crime_df
```

### 2.3.2 Resolving mismatch of state names

Some of the province names are mismatched between the crime rate csv file and the shpfile.

### 2.3.3 Performing relational join

```{r}
crime_df_mys <- mys_sf %>%
  left_join(crime_df,
            by = c("ADM1_EN" = "state")) %>%
  dplyr::select(ADM1_EN, ADM2_EN, category, type, crimes, year, Shape_Leng,  Shape_Area,  geometry)

crime_df_mys <- st_as_sf(crime_df_mys)
st_crs(crime_df_mys) <- st_crs(mys_sf)

crime_df_mys
```

### 2.3.4 Visualizing the distribution of crime

```{r}
#| eval: false

tmap_mode("plot")
choro <- tm_shape(crime_df_mys) +
  tm_fill("crimes", 
          style = "quantile", 
          palette = "Blues",
          title = "Crimes") +
  tm_layout(main.title = "Distribution of crime in West Malaysia",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2)

choro
```

# 3. Analysis

## 3.1 Exploratory Data Analysis (EDA)

## 3.1.1 Coorelation Plot

# 4. UI Design

## 4.1 Story Board

### 4.1.1 Prameters

### 4.1.2 Outputs

## 4.2 Layout

(Story Board)\[https://giniceseah.netlify.app/posts/2021-07-31-storyboard/\]

# 5. Reflection
