---
title: "In-class Exercise 3"
author: "Jeffery Lau"
date: 09-02-2024
date-modified: "last-modified"
description: |
  In this exercise we revisted on how to do proper CRS projection and transformations. Note: Last modified to update title only.
categories:
  - In-class
format:
  html:
    toc: true
execute: 
  eval: true
  echo: true
  warning: false
  freeze: false
---

# 1. Getting Started

The following code chunk uses p_load() to install and load the sf and tidyverse packages.

```{r}
pacman::p_load(tmap, sf, raster, spatstat, maptools, sp, tidyverse)
```

# 2. Importing data

The following code chunk will import the Master Plan data in ESRI shapefile and kml file format into R.

```{r}
mpsz_sf <- st_read(dsn = "data", layer = "MP14_SUBZONE_WEB_PL")
mpsz_sf <- st_set_crs(mpsz_sf, 3414)

sg_sf <- st_read(dsn = "data", layer = "CostalOutline")
sg_sf <- st_set_crs(sg_sf, 3414)

childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>%
  st_transform( crs = 3414)
childcare_sf <- st_set_crs(childcare_sf, 3414)

# st_union used to derive coastal outline sf data frame
sg_sf <- mpsz_sf %>% st_union()
tm_shape(sg_sf) + tm_borders()
```

Updated way of conversion is to use `as_owin()` and `as_ppp()` to convert sp dataframe Convert to Raster: spatstat use as(kd_childcareSG_adaptive, "SpatialGridDataFrame") if maptools:: is not installed

```{r}
acled_sf <- read_csv("./data/ACLED_Myanmar.csv") %>%
  st_as_sf(coords= c("longitude", "latitude"), crs=4326) %>%
  st_transform(crs = 32647) %>%
  mutate(event_date = dmy(event_date))

qtm(acled_sf)
```
